Tempus: Probabilistic Network Latency Verifier

### 2022-11-29 paper/main.tex
[Code] Finished evaluation

In our running example, we will assume that all edges in the topology will have $l_p$
and $l_q$ of exponential distribution with $\lambda = 1$, except $l_q(EZ)$ which will 
have a gamma distribution with $\alpha = 1$ and $\theta = 1$.

@@

In our example, we define $l_q(EZ)$ to be gamma distribution, which doesn't have a closed form convolution 
solution with the other latency distribution in the network, which is of exponential form.
Therefore, in order to convolve two arbitrary distributions, we need to leverage a numerical convolution 
method.

@@

In our running example, we could see that in Fig. \ref{fig:tree}, $\mathcal{E}_2$ and $\mathcal{E}_4$ share 
an identical path, $ADFZ$.
We could therefore memoize $cdf(\mathcal{L}_{ADFZ}, t)$ when we explore $\mathcal{{E}_2}$ and just bring the 
cached result when we explore $\mathcal{{E}_4}$.

With this optimization, we will do a chain convolution by the amount of path available in the network between 
the source and destination (in the case of Fig. \ref{fig:ex}, 3).

@@

In our running example, in the chain convolution process to get $\mathcal{L}_{ABEZ}$, we will analytically
convolve $l_p(AB)$, $l_q(AB)$, $l_p(BE)$, $l_q(BE)$, and $l_p(EZ)$ since they are of the same type (exponential) 
and then numerically convolve the resulting distribution with $l_q(EZ)$.
Fig. \ref{fig:grouping} illustrates how the chain convolution process will be broken down.

@@

\section{Evaluation}
\subsection{Scalability}

\subsubsection{States Explored}
\begin{figure}[h]
    \centering
    \includegraphics[scale=0.5]{states}
    \caption{States explored for temporal verification and functional verification}
    \label{fig:states}
\end{figure}

\subsubsection{Wall-clock Time}
\begin{figure}[h]
    \centering
    \includegraphics[scale=0.5]{wallclock}
    \caption{Time took for temporal verification}
    \label{fig:wallclock}
\end{figure}

@@

@@

\section{Future Works}

\section{Conclusion}

>>>>

In our running example, we will assume that all edges in the topology will have $l_p$
and $l_q$ of gamma distribution with $\lambda = 1$, except $l_q(EZ)$ which will 
have a lognormal distribution with $\alpha = 1$ and $\theta = 1$.

@@

In our example, we define $l_q(EZ)$ to be lognormal distribution, which doesn't have a closed form convolution 
solution with the other latency distribution in the network, which is of gamma distribution form.
Therefore, in order to convolve two arbitrary distributions, we need to leverage a numerical convolution 
method.

@@

In our running example, we could see that in Fig. \ref{fig:tree}, $\mathcal{E}_2$ and $\mathcal{E}_4$ share 
an identical path, $ADFZ$.
We could therefore memoize $cdf(\mathcal{L}_{ADFZ}, t)$ when we explore $\mathcal{E}_2$ and query the 
cached result when we explore $\mathcal{E}_4$.

With this optimization, we will do a chain convolution by the amount of path available in the network between 
the source and destination (which is 3 in the case of our running example in Fig. \ref{fig:ex}: $ACZ$, $ABEF$, 
and $ADFZ$).

@@

In our running example, in the chain convolution process to get $\mathcal{L}_{ABEZ}$, we will analytically
convolve $l_p(AB)$, $l_q(AB)$, $l_p(BE)$, $l_q(BE)$, and $l_p(EZ)$ since they are of the same type (gamma) 
and then numerically convolve the resulting distribution with $l_q(EZ)$.
Fig. \ref{fig:grouping} illustrates how the chain convolution process will be broken down.

@@

\section{Evaluation}

Spec
\begin{itemize}
    \item Julia
    \item i7-1065G7 16GB RAM
\end{itemize}

Network model
\begin{itemize}
    \item OSPF weight 1
    \item 0.1\% link failure rate
    \item $10^{-8}$ inaccuracy
    \item Reachability between two edge nodes
    \item Queuing latency: LogNormal distribution
    \item Link latency: either Gamma or Chisq (chosen at random)
    \item Bounded reachability with t=1000.0
\end{itemize}

\subsection{Runtime Scalability}

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.5]{scalability}
    \caption{Runtime performance of temporal verification, red dots represent the number of convolutions}
    \label{fig:scalability}
\end{figure}

\begin{itemize}
    \item The convolution step takes the most amount of time
    \item \textbf{Runtime performance correlates with the amount of convolutions} $->$ plot
    \item The amount of convolution is the result of the optimizations that we've done
\end{itemize}

\subsection{Equivalence Class Reduction}

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.5]{ec}
    \caption{Amount of temporal EC compared to functional EC, label represents ratio}
    \label{fig:ec}
\end{figure}

\begin{itemize}
    \item Compare the amount of temporal EC to the functional (unconsolidated) EC
    \item Corresponds to the consolidation strategy
    \item \textbf{More than 50\% reduction}
    \item Doesn't work equally on all topologies, depend on 
    \begin{itemize}
        \item How many functional EC are explored (for a given inaccuracy level)
        \item How many possible paths a given node pair have in those ECs
    \end{itemize}
\end{itemize}

@@

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.5]{optimization}
    \caption{Amount of convolution (in log scale) depending on what optimization strategy is applied}
    \label{fig:opt}
\end{figure}

\begin{itemize}
    \item While the amount of path in those states can explain the amount of EC reduction, to get to 
        the amount of convolution, we also need to consider the length of each of those path
    \item By considering the both amount of path and its length to counting the convolution, we could 
        compare consolidation and memoization strategy to the unoptimized version
    \item We could see that in the best case, the optimization strategies \textbf{reduce the amount of 
        convolution by 98\%} (in the case of GtsCe)
\end{itemize}

\subsection{Effect of Latency Distribution Type}

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.5]{distribution}
    \caption{The effect of distribution type on runtime performance of AttMpls, label represents the 
    performance in seconds and red dots represents the amount of \textit{numerical} convolution}
    \label{fig:dist}
\end{figure}

\begin{itemize}
    \item While the amount of convolution gave us a pretty good proxy for performance, we could see 
        that there's still some variations
    \item This is due to the fact that we used two kinds of convolutions: analytical and numerical.
    \item Since numerical convolution is slower, we could see that for the same topology and the same amount of 
        overall convolution, \textbf{more numerical convolution will result in longer runtime performance}.
    \item In the worst case (all numerical), the grouping optimization would have no effect
\end{itemize}

@@

\section{Future Works}
\begin{itemize}
    \item Addressing independence: convolution treats each probability distribution as independent 
        with each other. Latency verifier that relaxes this assumption (or a better encoding) would be 
        interesting future works
    \item Optimizing functional verification: a functional verifiers that produces functional ECs that 
        has already been consolidated would be interesting future works
    \item Extracting Latency Information from Measurement Data: While Tempus serve as a framework for 
        latency verification, future work needs to be done in translating measurement data into latency 
        distributions. Julia supports distribution fitting given a set of data. 
\end{itemize}

\section{Conclusion}
\begin{itemize}
    \item Tempus is a scalable latency SLA verifier. It could probabilistically verify in the order of minutes
    \item Consolidation and memoization is an effective optimization strategy, although the exact magnitude of 
        its effectiveness will depend on the network topology, specifically on the amount of paths and their 
        length.
    \item The grouping optimization strategy is effective only if most distributions in the network can be 
        convolved analytically. In the worst case, if most of them can't (like LogNormal only) then this 
        optimization would have no improvement.
\end{itemize}
